---
title: "Introduction to ggplot2"
---

## Session-1 :Introduction to ggplot2
Firstly load the data from the github repository and check whehter the data has been imported correctly. To do so, you need to call "RCurl" library.

```{r}
library(RCurl)
xdat <- read.csv(text=getURL("https://raw.githubusercontent.com/jaynal83/MDVR/master/data/anscombe_quartet_modified.csv"), header = T, sep = ",")
head(xdat)
```


# step-1: Define the plot and map the aesthetics
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
plot1 <- ggplot(data = xdat, aes(x=x, y=y))
print(plot1)
```

# step-2: Adding facet (panel for each group)
```{r}
plot1 <- plot1 + facet_wrap(~group)
print(plot1)
```

# step-3: Adding points to the plot
```{r}
plot1 <- plot1 + geom_point()
print(plot1)
```

# step-4: Modifying colour, shape, and size of the points
```{r}
plot1 <- plot1 + geom_point(size = 5, fill = "red", color = "darkgreen", shape = 21)
print(plot1)
```

# step-5: Adding fitted line (linear fit)
```{r}
plot1 <- plot1 + geom_smooth(method = "lm")
print(plot1)
```

# step-6: removing confidence interval and modify line type, size and colour
```{r}
plot1 <- plot1 + geom_smooth(method = "lm", se = F, linetype = 1, size = 2, col = "orange")
print(plot1)
```


# step-7: Removing background colour
```{r}
plot1 <- plot1 + theme_bw()
print(plot1)
```

# step-8: Adding plot title and subtitle
```{r}
plot1 <- plot1 + ggtitle(
  label = "Anscombe Quartet", 
  subtitle = "Scatter plot"
  ) 
print(plot1)
```


# step-9: Modifying size, colour and position of plot title
```{r}
plot1 <- plot1 + theme(
    plot.title = element_text(
      size = 25, 
      face = "bold", 
      color = "orange", 
      hjust = 0.5),
    plot.subtitle = element_text(
      size = 15, 
      face = "italic", 
      colour = "darkgreen", 
      hjust = 0.5)
  )
print(plot1)
```

